#Description Given is Walmart's transactional dataset. The task is to classify customer shopping trips The goal of this homework is to create a model which results will be within the top 150 from the Kaggle leaderboard:

```{r}
library(caret)
library(dplyr)
library(tidyr)
library(randomForest)

data <- read.csv(("/Users/Ashmitha/Downloads/train.csv"), header=TRUE, sep=",", encoding="UTF-8")
head(data)

```

TripType - a categorical id representing the type of shopping trip the customer made. This is the ground truth that you are predicting. TripType_999 is an "other" category.

VisitNumber - an id corresponding to a single trip by a single customer. It is non-predictive hence will be excluded.

Weekday - the weekday of the trip.

Upc - the UPC number of the product purchased.

ScanCount - the number of the given item that was purchased. A negative value indicates a product return.

DepartmentDescription - a high-level description of the item's department.

FinelineNumber - a more refined category for each of the products, created by Walmart.

data %>% filter(VisitNumber == 7)

#Extracting features
ProductsCountInVis - Products purchased per visit number where scan count is 0 or more

newDepartmentsInVis = data %>%
  group_by(VisitNumber) %>%
  summarise(
    ProductsCountInVis = sum(ScanCount >=0)
  )

head(newDepartmentsInVis)

ProductsCountInDept - Products purchased per visit number and DepartmentDescription where scan count is 0 or more

newDepartments = data %>%
  group_by(VisitNumber, DepartmentDescription) %>%
  summarise(
    ProductsCountInDept = sum(ScanCount >=0)
  )
head(newDepartments)

Join on newDepartmentsInVis and newDepartments on VisitNumber
tempData <- merge(newDepartmentsInVis,newDepartments,by="VisitNumber")
head(tempData)

tempData$ProductsCountInDept = tempData$ProductsCountInDept / tempData$ProductsCountInVis  

#Data Tidying
Spread DepartmentDescription, across multiple columns(VisitNumber,ProductsCountInVis), missing values will be replaced with fill value which is 0.

library(tidyr)

TransformedDepartment = tempData %>%
  spread(DepartmentDescription, ProductsCountInDept, fill=0)

head(TransformedDepartment)

newData is a composition of data per VisitNumber on a particular day the count of products sold, with number of returned products.

newData = data %>%
  group_by(VisitNumber) %>%
  summarise(
    TripType = head(TripType, 1),
    day = unique(Weekday)[1], 
    ProductsCount = sum(ScanCount >= 0),
    unique_FinelineNumber = length(unique(FinelineNumber)),
    no_of_returned_products = sum(ScanCount < 0),
    unique_departments = length(unique(DepartmentDescription)),
    sum_of_products = sum(ScanCount)
  )

head(newData)

Joining newData with TransformedDepartment on VisitNumber, by this the products count on a particular day for each department is shown.
finalData <- merge(newData,TransformedDepartment,by="VisitNumber")
head(finalData)

colnames(finalData) = make.names(colnames(finalData))

finalData$TripType <-factor(finalData$TripType)
summary(finalData)

# dividing dataset into train and test part
indxSample <- sample(seq_len(nrow(finalData)), size = 1000)
sample <- finalData[indxSample, ]

prop.table(table(sample$TripType))

nzv <- nearZeroVar(sample)
names(sample)[nzv]
data <- sample[, -nzv]

data$TripType <- factor(data$TripType)
levels(data$TripType)

names(data) = make.names(names(data), unique = FALSE)
sample=data
set.seed(1313)
indxTrain <- createDataPartition(y = finalData$TripType, p = 0.75)
str(indxTrain)

kagTrain <- finalData[indxTrain$Resample1,]
kagTest <- finalData[-indxTrain$Resample1,]

library(randomForest)

# using Random Forest classifier
rf <- randomForest(TripType~., data=kagTrain)
confusionMatrix = confusionMatrix(kagTest$TripType, predict(rf, kagTest))$overall
scores <- predict(rf, kagTest, type = "prob")

# calculating final score
myScores <- sapply(1:nrow(kagTest), function(i){
  scores[i, kagTest$TripType[i]]
})

mean(-log(pmax(myScores,0.05)))

#using important features
```{r}
ffit <- randomForest(factor(TripType) ~ sum_of_products + unique_departments +VisitNumber, data=kagTrain, importance = TRUE)
print(ffit)

scores <- predict(ffit, kagTest, type="prob")
head(scores)

myScores <- sapply(1:nrow(kagTest), function(i){
  scores[i, as.character(kagTest$TripType[i])]
})

mean(-log(pmax(myScores,0.05)))
```

#KNN 
```{r}
knnFit <- knn3(TripType ~ ., data = kagTrain, k=20, prob=TRUE)
scoresknn <- predict(knnFit, kagTest, type="prob")
head(scoresknn)

myScoresknn <- sapply(1:nrow(kagTest), function(i){
  scoresknn[i, as.character(kagTest$TripType[i])]
})

mean(-log(pmax(myScoresknn,0.05)))
```

#KNN with important features
```{r}
knnFit <- knn3(TripType ~ sum_of_products + unique_departments +VisitNumber, data = kagTrain, k=20, prob=TRUE)
scoresknn1 <- predict(knnFit, kagTest, type="prob")
head(scoresknn1)

myScores1 <- sapply(1:nrow(kagTest), function(i){
  scoresknn1[i, as.character(kagTest$TripType[i])]
})

mean(-log(pmax(myScores1,0.05)))
```
